Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-01-27 analysis/paper_draft3.Rmd
small edits to intro and discussion

The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data. A commonly adopted practice is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho2021, 5252; @Ibrahim2015, 3734]. Specifically, lower values (closer to 0) of $\alpha$ should be selected if incompatibility is expected to be high. Higher values (closer to 1) of the power prior parameter should be selected if incompatibility is expected to be low. Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately corresponding to the likely degree of incompatibility between the current and historical strike data in the particular case. Since this overall procedure is to estimate bias to support the bringing of a _Batson_ challenge, the senstivity of bias detection to the value of $\alpha$ is used to evaluate sensitivity of the estimation of bias, not to indicate whether a _Batson_ challenge should succeed.

@@

In future work we hope to examine data-driven approach to selecting $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data. Knowing how to quantify incompatibility and how to specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022]

>>>>

The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data. One approach is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho2021, 5252; @Ibrahim2015, 3734]. Specifically, lower values (closer to 0) of $\alpha$ should be selected if incompatibility is expected to be high. Higher values (closer to 1) of the power prior parameter should be selected if incompatibility is expected to be low. Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately corresponding to the likely degree of incompatibility between the current and historical strike data in the particular case. Since this overall procedure is to estimate bias to support the bringing of a _Batson_ challenge, the senstivity of bias detection to the value of $\alpha$ is used to evaluate sensitivity of the estimation of bias, not to indicate whether a _Batson_ challenge should succeed.

@@

In future work we hope to examine data-driven approaches to selecting $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data. Knowing how to quantify incompatibility and how to specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022]. In any case, any formulation for estimating $\alpha$ based on incompabitiliy between current and historical trials would require justification in court (or by practitioners).
